This thesis presents TF-Rehancer V5, a time-frequency speech enhancement architecture for direct full-band processing. The model treats high-frequency components in 48 kHz speech as observed but noisy spectral evidence rather than as missing information to be generated from a band-limited input. TF-Rehancer uses sampling-frequency-independent STFT analysis, power-law compressed complex spectral conditioning, low-band encoder-guided refinement of observed high-band query features, and local complex time-frequency filtering over the original noisy STFT.

The evaluation is organized around two roles. First, a 16 kHz VoiceBank+DEMAND benchmark is used to compare quality and efficiency against lightweight speech enhancement references. In this setting, TF-Rehancer achieves PESQ-WB 3.334, DNSMOS SIG 3.405, DNSMOS BAK 4.042, and DNSMOS OVRL 3.129 with 0.470M parameters and 1.890G model-side MACs. Compared with the locally evaluated FastEnhancer-M checkpoint, TF-Rehancer improves PESQ-WB and DNSMOS SIG/BAK/OVRL with lower MACs, while FastEnhancer-M remains stronger in DNSMOS P.808, SCOREQ, SI-SDR, STOI, and ESTOI. The result therefore supports a competitive quality-efficiency trade-off rather than uniform dominance.

Second, 48 kHz experiments are used as controlled full-band ablations. Models are trained with VCTK clean speech mixed with DNS full-band noise and evaluated on the VoiceBank+DEMAND 48 kHz test set. These experiments are not used as external full-band system rankings. The 48 kHz ablations show that replacing complex TF filtering with direct spectrum mapping degrades full-reference, spectral, and non-intrusive metrics. They also show that decoder cross-attention changes the trade-off between perceptual scores and spectral fidelity, and that a 5 kHz low-band cutoff provides a balanced operating point under the controlled ablation protocol.
